Optimal Pricing and Informal Sharing: \\ Evidence from Piped Water in Manila

### 2018-09-23 phil_paper/jmp_wjv_v2.tex
fixing some graph stuff

\documentclass[12pt]{article}
\usepackage[margin=1in] {geometry}                
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
\pagestyle{plain}                           % Eliminates both the header and footer
\usepackage[parfill]{parskip}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{array}
%\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{tkz-euclide}
\usepackage{longtable}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{verbatim}  %so that you can comment blocks of code
\usepackage{setspace}  %change the spacing
\usepackage[authoryear]{natbib}
\usepackage{bbm}
\usetikzlibrary{calc}
%\usepackage{csvsimple}
\usepackage[toc,page]{appendix}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{authblk}
\usepackage[bottom]{footmisc}
\usepackage{titling}
%% \newcommand{\KEYtitle}{Even Weights}
%\usepackage[nomarkers]{endfloat}
%\bibliography{library} 
%\renewcommand{\efloatseparator}{\mbox{}}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage[export]{adjustbox}
\usetikzlibrary{positioning}

@@

Conventional pricing policies often rely on the assumption that each water connection serves a single household; however, many households in developing cities share with neighbors instead of purchasing their own connections.  For example, 70\% of African cities report frequent water sharing (\cite{keener2010provision}) ranging from fetching from neighbors with hoses and buckets in Tangiers, Morocco (\cite{devoto2012happiness}) to drawing from single taps that each serve compounds in Kumasi, Ghana (\cite{whittington1992possible}).\footnote{\cite{nauges2006water} finds further evidence of widespread water sharing in Sri Lanka.}  In Manila, I establish that 23\% of households access water through a neighbor's connection by connecting plumbing directly to the meter (\input{tables/access_pipes}\unskip\%), extending plastic pipes (\input{tables/access_single_tap}\unskip\%), or fetching water with containers (\input{tables/access_fetch}\unskip\%).  While previous studies have focused on particular suburbs, this paper is the first to my knowledge to measure the role of water sharing in extending access across an urban population.

@@

Combining the model and estimates, I simulate a series of counterfactual policies.  First, I consider a connection fee subsidy policy consistent with recommendations from the World Bank, NGOs, and other researchers to improve water access for the poor (\cite{komives2005water,wsup,jimenez2014factors,mcintosh2003asian}).  I discount the fixed fee by 12\% while increasing the tariff schedule proportionally to ensuring that the water provider exactly covers costs.  This policy reduces access by incentivizing households to substitute away from sharing arrangements toward both individual connections and water vendors (which mainly include deepwells and water refilling stations).

@@

Finally, I compare the optimal tariff with a social tariff that maximizes consumer surplus only for the bottom 50\% of users and is intended to capture goals of ensuring minimum usage and access as well as income redistribution.  The social tariff features a much higher fixed fee and lower marginal price, does not improve access to piped water, and reallocates little consumer surplus to low-demand households at a large cost to average welfare.  By contrast, in a counterfactual world without sharing, the social tariff features a fixed fee subsidy and increasing marginal prices; it also increases access and efficiently redistributes consumer surplus to low users as compared the optimal tariff.  These results provide evidence that the presence of informal sharing networks constrains pricing policy as an effective tool for redistribution.

@@

%Monthly quantity choices across different households and at different prices reveal information about household demand.  Observing multiple quantities choices for the same household allows for a rich characterization of demand heterogeneity at the household-level.  
Evaluating different pricing policies requires measuring demand at the household level as well as costs associated with providing piped water.  For information about these attributes, I developed a partnership with one of the regulated, private providers in Manila who provided both billing records as well as information about the regulatory structure, pricing, and production costs.  Complete billing records for each water connection including the monthly meter reading, the total bill charged, and the precise location for each meter serve as the primary measure of quantity choices.  These data span 2010 to 2015, during which connections grew from 800,000 to 1.5 million water connections, generating nearly 100 million transaction records.  Connections are divided into residential (90\%), semi-business (4\%), commercial (5\%), and industrial (1\%).  This paper focuses on residential and semi-business connections.\footnote{Semi-business connections are largely composed of households that own small roadside stands.  These connections provide a useful source of price variation which is described in Appendix~\ref{appendix:ratechangeelasticity}.}  The monthly meter reading is used as the measure of consumption for each connection.  Meter readings are rounded to the nearest cubic meter, and I limit consideration to meter readings below 120 $\text{m}^{3}$ in order to exclude pipe leaks as well as meter reading errors (readings above 120 $\text{m}^{3}$ compose less than 0.5\% of the sample).

@@

A full analysis of water usage in Manila also requires describing households that choose to use water from local vendors.  For information on households that are using water from alternative sources, I turn to the 2010 Census of Population and Housing.  Households are defined as connected to the piped network if they report using a ``faucet community water system'' for their cooking, laundry, or bathing needs.\footnote{The census was designed largely for rural populations so it does not directly ask whether households are connected to a piped water network.}  Alternatively, I define households as using from a water vendor if they report any other category, which includes deep wells, peddlers, or other sources.  I then construct similar demographic measures to the connection survey using questions in the census and compare these measures across these data sources in Table~\ref{table:pawstocensus} of Appendix~\ref{appendix:pawstocensus}.  Households that own a connection in the connection survey are similar to households connected to the piped water network in the Census data.  To construct a profile of water source choices for the full population, I randomly sample from the pool of households using from vendors in 8 regions of the city so that the data form a representative sample when combined with households covered by the connection survey in each small area.\footnote{Regions are chosen to overlap with ward administrative/political boundaries while ensuring that I am able to obtain a large enough sample of vendor households for each area to perform the counterfactual analysis in Section~\ref{section:fixedcostestimation}.}

@@

The water provider also maintains call records of all service complaints, which can be linked to individual connections.  This data is used to measure instances where households face unanticipated shocks to their water source as part of the identification strategy to estimate the hassle costs of sharing water.  Finally, the water provider tracks data on the marginal cost of pumping an additional cubic meter of water as well as the costs associated with installing new connections.  These data provide inputs into the provider's budget constraint, which are necessary for computing counterfactual pricing regimes.  Supplemental survey data from a subset of Manila include income measures that are used to impute income for the remainder of the sample across a variety of demographic variables and their interactions.

@@

\begin{table}
\centering
\captionsetup{justification=centering}
\caption{ Water Connections  by Number of Households Served}\label{descriptives}
\input{tables/new_descriptives} \\
\footnotesize{The data include billing records merged with the connection survey. \\ Income is imputed using location, demographic variables, and their interactions.}
\end{table}

Table~\ref{descriptives} also summarizes water use and sharing behavior for these connections, merging information from the connection survey with the billing records.  Larger numbers of households sharing a water connection are correlated with higher total water usage, but less water per person.\footnote{Total People Served by a connection includes both the owning household and any other households using a connection.}  This decline in usage per person may have many possible drivers: (1) sharing households face higher prices by locating at a higher point on the tariff schedule (Figure~\ref{figure:tariff}), (2) sharing households may face hassle costs in fetching water from neighboring taps, (3) these price effects may incentivize low-demand households to select into sharing relationships, consistent with the demographic patterns, and (4) low-demand households may have idiosyncratic preferences for sharing relationships.  On average, monthly water bills (inclusive of service fees) take up around \input{tables/incshr}\unskip\% of household income, which does not include other unobserved costs in accessing water.  These findings suggest that policies which affect sharing choices can have significant welfare consequences.

@@

Sharing households must negotiate over how to pay the monthly fixed fees.  The model assumes that households commit to dividing any fixed fees at the start of the sharing arrangement.  Under quasi-linear utility, any fixed transfers are exogenous to monthly consumption decisions.  For ease of exposition, this section assigns a fixed fee, $F_i$ to each household and postpones discussion of how the total fixed fee may be divided between households to Section~\ref{section:watersourcechoice}.

@@

6\unskip\% of households source their water from small-scale water vendors.  These sources are collapsed into a single vendor category, which is modeled as a perfect substitute to piped water and is characterized by a monthly fixed fee, $F_v$, and a marginal price, $p_v$.  The monthly fixed fee, $F_v$, is assumed to capture primarily fixed differences in quality of service between vended water and other sources.  Similarly, the marginal price, $p_v$, includes both the price per unit as well as any differences in water quality or ease of use relative to piped water.  Demand for this source becomes a simple linear function of prices given by equation (\ref{equation:vendordemand}).
\begin{align}\label{equation:vendordemand}
w^{*} = \Gamma_i - \alpha_i p_v
\end{align}
Since the data do not record monthly consumption information for households using from vendors, these households do not contribute to the likelihood function above.

@@

\begin{table}
\centering
\caption{Demand Estimates}\label{table:demandestimates}
\input{tables/step_1_estimates_group} \\
\vspace{.1cm}
{\footnotesize{Total Observations:  349,265   Total Connections:  7,151} }
\end{table}
To interpret these estimates in economic terms, I compute the elasticity of demand with respect to price at the connection-level.  Given a linear demand structure, the elasticity of demand depends on both the price sensitivity and fixed preferences for water.  As a simplified example, consider the calculation for elasticity under a single marginal price and linear demand specification:
\begin{align*}
\frac{\delta Q}{\delta P} \frac{P}{Q} = \alpha_{C} \frac{P}{\gamma_{C}-\alpha_{C} P}
\end{align*}
According to this expression, increasing the fixed preference for water consumption, $\gamma_C$, while holding constant the price sensitivity, $\alpha_C$, mechanically reduces the elasticity.  Many previous studies estimate heterogeneity in $\gamma_C$ alongside a single $\alpha_C$ parameter for the population (\cite{mcrae2014infrastructure}, \cite{olmstead2009reduced}, \cite{szabo2015value}).  This paper builds on these previous studies by estimating heterogeneity in both terms to allow to for a broader range of possibilities, including that larger users may be relatively more sensitive to prices than smaller users.  To compare estimates in this paper with previous papers under non-linear pricing schedules, this approach follows the same formula for computing elasticities by (1) taking draws of $\epsilon$ and $\eta$, (2) computing predicted demand under prices that are 1\% greater than current prices, and (3) comparing the average percentage change in demand before and after the price change.  For tariff blocks equal to zero, prices are increased from 0 to 1 PhP following \cite{szabo2015value}.

@@

\begin{table}
\centering
\caption{Elasticity Estimates}\label{table:elasticity}
\input{tables/elasticity_estimates} \\
\vspace{.5cm}
{\footnotesize Total Observations:  349,265   Total Connections:  7,151 \\
Standard errors are bootstrapped at the connection level.}
\end{table}

The demand estimation yields an average elasticity of -\input{tables/mean_elasticity}\unskip, which can be interpreted as the percentage change in quantity due to a one percentage change in price.  This estimate also closely matches reduced-form elasticity estimates from price-change quasi-experiment detailed in Appendix~\ref{appendix:ratechangeelasticity}.  This estimate is well below the -0.98 estimate of \cite{szabo2015value} but in the range of other similar studies in developing world that find elasticities between -0.01 and -0.81 (\cite{diakite2009proposal}, \cite{strand2005water}).

@@

Figure~\ref{figure:predictedwateruse} compares the predicted histogram of consumption with observed values.  Vertical red lines indicate the price thresholds in the tariff schedule.  By roughly matching the smoothness of the distributions especially across price thresholds, this model improves on standard models in the literature, which often predict excessive bunching at thresholds (\cite{mcrae2014infrastructure}, \cite{olmstead2009reduced}, \cite{szabo2015value}).  The model has difficultly rationalizing low levels of consumption likely because households face zero prices below this threshold.  Since preference and consumption shocks are assumed to be normally distributed, the model predicts some negative consumption values.  In the counterfactual exercises, I round these negative realizations to zero consumption.  Likewise, the model also cannot account for bunching at zero (likely capturing households that go on vacation).  The model also predicts greater bunching around the upper price threshold despite little observed bunching in the data as large users tend to ignore this smaller jump in price.

@@

\begin{table}
\centering
\caption{Correlation Estimates}\label{table:correstimates}
\input{tables/corr_table_group}
%\vspace{.3cm}
%\footnotesize{1 indexes correlations with first movers \\ while 2 indexes correlations with second movers}
\end{table}

@@

\begin{table}
\centering
\caption{Hassle Cost Estimates}\label{table:step2estimates}
\input{tables/step_2_estimates} \\
\vspace{.2cm} 
\footnotesize{Leaking Connections:  548,   Neighboring Connections:  2,151,   Total Obs:  20,299  \\ Standard errors are bootstrapped at the leaking connection-neighborhood level. }
\end{table}

@@

\begin{table}
\centering
\caption{Simulated Method of Moments Estimates}\label{table:smmestimates}
\input{tables/step_3_estimates_group}\\
\vspace{.05cm}
\footnotesize{ Total Households:  \input{tables/totalhhsmm_group}   Total Wards:  599 \\ Standard errors are bootstrapped at the connection level.   }
\end{table}

@@

\begin{table}
\centering
\caption{Connection Fee Discount: Out-of-Sample Test Results}\label{table:oosestimates}
\input{tables/oos_discount_estimates_group} \\ 
\footnotesize{Consumer surplus is measured in PhP/month.}
\end{table}

@@

Table~\ref{table:fixedfeediscount} summarizes the impacts of this change on source choices and consumer welfare measured in terms of the difference between current utility and utility from using the outside vendor option as in Section~\ref{section:oostest}.  While the increase in the tariff drives large users to substitute toward vendor options, the decrease in the fixed price allows some small users to join together and purchase shared connections, resulting in no change in the share of households using from a neighbor.  The third row indicates even greater substitution toward water vendors of around \input{tables/discount_vendor_change_group}percentage points in response to the higher tariff.  These effects combine to reduce average consumer surplus by \input{tables/discount_avg_surplus_change_group}PhP per month, which is over half of the initial fixed fee discount of \input{tables/c_fee_discount_group}PhP.  This decline translates into a \input{tables/discount_avg_welfare_change_group}percent loss of monthly household income.  Households in the bottom \input{tables/group_percentile_group}\unskip th percentile of water users also experience lower consumer surplus although the magnitude is much less, amounting to \input{tables/discount_low_welfare_change_group}percent of monthly household income.  Taken together, these results suggest that connection fee discount policies in this context can have unintended distributional consequences by reducing both access to piped water and welfare for the intended beneficiaries. %\footnote{Consumer surplus is measured relative to utility from using vended water.}  
% a drop of \input{tables/discount_sharing_change_group}percentage points in households using water from their neighbors The decrease in the fixed fee and increase in the tariff both act to decrease the incentives for sharing a connection and leading to no change in households using water from their neighbors. 
\begin{table}
	\centering
	\caption{Fixed Fee Discount Counterfactual}\label{table:fixedfeediscount}
	\input{tables/discount_estimates_group} \\
	\footnotesize{Low Users include the bottom 50\% of water users. \\
	Surplus is measured in terms of PhP/month.}
\end{table}

@@

\begin{table}[!hbp]
\centering
\caption{Optimal Tariffs Without Sharing}\label{table:optimaltariffnosharing}
\resizebox{\textwidth}{!}{%
\input{tables/welfare_no_sharing_groups}
}	\\
\footnotesize{$\text{ }^{*}$ The First-Best fee does not include individual transfers imposed by the regulator. \\  Social tariffs place all welfare weight on the bottom 50\% of water users. \\ Surplus is measured in terms of PhP/month.}
\end{table}

>>>>

\documentclass[12pt]{article}
\usepackage[margin=1in] {geometry}                
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
\pagestyle{plain}                           % Eliminates both the header and footer
\usepackage[parfill]{parskip}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{array}
\usepackage{threeparttable}
%\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{tkz-euclide}
\usepackage{longtable}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{verbatim}  %so that you can comment blocks of code
\usepackage{setspace}  %change the spacing
\usepackage[authoryear]{natbib}
\usepackage{bbm}
\usetikzlibrary{calc}
%\usepackage{csvsimple}
\usepackage[toc,page]{appendix}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{authblk}
\usepackage[bottom]{footmisc}
\usepackage{titling}
%% \newcommand{\KEYtitle}{Even Weights}
%\usepackage[nomarkers]{endfloat}
%\bibliography{library} 
%\renewcommand{\efloatseparator}{\mbox{}}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage[export]{adjustbox}
\usetikzlibrary{positioning}

@@

Conventional pricing policies often rely on the assumption that each water connection serves a single household; however, many households in developing cities share with neighbors instead of purchasing their own connections.  For example, 70\% of African cities report frequent water sharing (\cite{keener2010provision}) ranging from fetching from neighbors with hoses and buckets in Tangiers, Morocco (\cite{devoto2012happiness}) to drawing from single taps that each serve compounds in Kumasi, Ghana (\cite{whittington1992possible}).\footnote{\cite{nauges2006water} finds further evidence of widespread water sharing in Sri Lanka.}  In Manila, I establish that 23\% of households access water through a neighbor's connection by connecting plumbing directly to the meter (\input{tables/access_pipes}\unskip\%), extending plastic pipes (\input{tables/access_single_tap}\unskip\%), or fetching water with containers (\input{tables/access_fetch}\unskip\%).  While previous studies have focused on specific suburbs, this paper is the first to my knowledge to measure the role of water sharing in extending access across an urban population.

@@

Combining the model and estimates, I simulate a series of counterfactual policies.  First, I consider a connection fee subsidy policy consistent with recommendations from the World Bank, NGOs, and other researchers to improve water access for the poor (\cite{komives2005water,wsup,jimenez2014factors,mcintosh2003asian}).  I discount the fixed fee by 12\% while increasing the tariff schedule proportionally to ensuring that the water provider exactly covers costs.  This policy reduces access by incentivizing households to substitute toward water vendors (which mainly include deepwells and water refilling stations).

%substitute away from sharing arrangements toward both individual connections and
%This policy reduces access by incentivizing households to substitute away from sharing arrangements toward both individual connections and water vendors (which mainly include deepwells and water refilling stations).

@@

Finally, I compare the optimal tariff with a social tariff that maximizes consumer surplus only for the bottom 50\% of users and is intended to capture goals of ensuring minimum usage and access as well as income redistribution.  With sharing, social tariffs are ineffective at reallocating surplus from large to small users; however in a counterfactual without sharing, social tariffs achieve substantial redistribution, over doubling consumer surplus for small users.  This evidence suggests that the presence of informal sharing networks constrains non-linear pricing as an effective tool for redistribution.

%features a much lower fixed fee and higher marginal price and redistributes substantial surplus to low users as compared to the optimal tariff.

%features a slightly lower fixed fee and higher marginal price, produces universal access to piped water and reduces sharing, but reallocates little consumer surplus to low-demand households.

%and increasing marginal prices; it also increases access and efficiently redistributes consumer surplus to low users as compared the optimal tariff.  These results provide evidence that the presence of informal sharing networks constrains pricing policy as an effective tool for redistribution.

%The social tariff features a much higher fixed fee and lower marginal price, does not improve access to piped water, and reallocates little consumer surplus to low-demand households at a large cost to average welfare.  By contrast, in a counterfactual world without sharing, the social tariff features a fixed fee subsidy and increasing marginal prices; it also increases access and efficiently redistributes consumer surplus to low users as compared the optimal tariff.  These results provide evidence that the presence of informal sharing networks constrains pricing policy as an effective tool for redistribution.

@@

%Monthly quantity choices across different households and at different prices reveal information about household demand.  Observing multiple quantities choices for the same household allows for a rich characterization of demand heterogeneity at the household-level.  
Evaluating different pricing policies requires measuring demand at the household level as well as costs associated with providing piped water.  For information about these attributes, I developed a partnership with one of the regulated, private providers in Manila who provided both billing records as well as information about the regulatory structure, pricing, and production costs.  Complete billing records for each water connection including the monthly meter reading, the total bill charged, and the precise location for each meter serve as the primary measure of quantity choices.  These data span 2010 to 2015, during which connections grew from 900,000 to 1.5 million water connections, generating nearly 100 million transaction records.  Connections are divided into residential (90\%), semi-business (4\%), commercial (5\%), and industrial (1\%).  This paper focuses on residential and semi-business connections.\footnote{Semi-business connections are largely composed of households that own small roadside stands.  These connections provide a useful source of price variation which is described in Appendix~\ref{appendix:ratechangeelasticity}.}  The monthly meter reading is used as the measure of consumption for each connection.  Meter readings are rounded to the nearest cubic meter, and I limit consideration to meter readings below 100 $\text{m}^{3}$ in order to exclude pipe leaks as well as meter reading errors (readings above 100 $\text{m}^{3}$ compose less than 0.7\% of the sample).

@@

A full analysis of water usage in Manila also requires describing households that choose to use water from local vendors.  For information on households that are using water from alternative sources, I turn to the 2010 Census of Population and Housing.  Households are defined as connected to the piped network if they report using a ``faucet community water system'' for their cooking, laundry, or bathing needs.\footnote{The census was designed largely for rural populations so it does not directly ask whether households are connected to a piped water network.}  Alternatively, I define households as using from a water vendor if they report any other category, which includes deep wells, peddlers, or other sources.  I then construct similar demographic measures to the connection survey using questions in the census and compare these measures across these data sources in Table~\ref{table:pawstocensus} of Appendix~\ref{appendix:pawstocensus}.  Households that own a connection in the connection survey are similar to households connected to the piped water network in the Census data.  To construct a profile of water source choices for the full population, I randomly sample from the pool of households using from vendors in 599 wards of the city so that the data form a representative sample when combined with households covered by the connection survey in each small area.
%\footnote{Regions are chosen to overlap with ward administrative/political boundaries while ensuring that I am able to obtain a large enough sample of vendor households for each area to perform the counterfactual analysis in Section~\ref{section:fixedcostestimation}.}

@@

The water provider also maintains call records of all service complaints, which can be linked to individual connections.  This data is used to measure instances where households face unanticipated shocks to their water source as part of the identification strategy to estimate the marginal costs of sharing water.  Finally, the water provider tracks data on the marginal cost of pumping an additional cubic meter of water as well as the costs associated with installing new connections.  These data provide inputs into the provider's budget constraint, which are necessary for computing counterfactual pricing regimes.  Supplemental survey data from a subset of Manila include income measures that are used to impute income for the remainder of the sample across a variety of demographic variables and their interactions.

@@

\begin{table}
\centering
%\captionsetup{justification=centering}
\begin{threeparttable}
\caption{ Water Connections  by Number of Households Served}\label{descriptives}
\input{tables/new_descriptives}
\begin{tablenotes}
\item \footnotesize{The data include billing records merged with the connection survey.  Income is imputed using location, demographic variables, and their interactions.}
\end{tablenotes}
\end{threeparttable}
\end{table}

Table~\ref{descriptives} summarizes water use and sharing behavior for these connections, merging information from the connection survey with the billing records.  Larger numbers of households sharing a water connection are correlated with higher total water usage, but less water per person.\footnote{Total People Served by a connection includes both the owning household and any other households using a connection.}  This decline in usage per person may have many possible drivers: (1) sharing households face higher prices by locating at a higher point on the tariff schedule (Figure~\ref{figure:tariff}), (2) sharing households may face hassle costs in fetching water from neighboring taps, (3) these price effects may incentivize low-demand households to select into sharing relationships, consistent with the demographic patterns, and (4) low-demand households may have idiosyncratic preferences for sharing relationships.  On average, monthly water bills (inclusive of service fees) take up around \input{tables/incshr}\unskip\% of household income, which does not include other unobserved costs in accessing water.  These findings suggest that policies which affect sharing choices can have significant welfare consequences.

@@

Sharing households must negotiate over how to pay the monthly fixed costs (including the fixed fee paid to the firm and any unobserved costs of using a connection).  The model assumes that households commit to dividing any fixed costs at the start of the sharing arrangement.  Under quasi-linear utility, any fixed transfers are exogenous to monthly consumption decisions.  For ease of exposition, this section assigns a fixed cost, $F_i$ to each household and postpones discussion of how the total fixed cost may be divided between households to Section~\ref{section:watersourcechoice}.

@@

6\unskip\% of households source their water from small-scale water vendors.  These sources are collapsed into a single vendor category, which is modeled as a perfect substitute to piped water and is characterized by a monthly fixed cost, $F_v$, and a marginal price, $p_v$.  The monthly fixed cost, $F_v$, is assumed to capture primarily fixed differences in quality of service between vended water and other sources.  Similarly, the marginal price, $p_v$, includes both the price per unit as well as any differences in water quality or ease of use relative to piped water.  Demand for this source becomes a simple linear function of prices given by equation (\ref{equation:vendordemand}).
\begin{align}\label{equation:vendordemand}
w^{*} = \Gamma_i - \alpha_i p_v
\end{align}
Since the data do not record monthly consumption information for households using from vendors, these households do not contribute to the likelihood function above.

@@

\begin{table}
\centering
\caption{Demand Estimates}\label{table:demandestimates}
\input{tables/step_1_estimates_group} \\
\vspace{.2cm}
{\footnotesize{Sample includes 349,265 billing months for 7,151 connections. } }
\end{table}
To interpret these estimates in economic terms, I compute the elasticity of demand with respect to price at the connection-level.  Given a linear demand structure, the elasticity of demand depends on both the price sensitivity and fixed preferences for water.  As a simplified example, consider the calculation for elasticity under a single marginal price and linear demand specification:
\begin{align*}
\frac{\delta Q}{\delta P} \frac{P}{Q} = \alpha_{C} \frac{P}{\gamma_{C}-\alpha_{C} P}
\end{align*}
According to this expression, increasing the fixed preference for water consumption, $\gamma_C$, while holding constant the price sensitivity, $\alpha_C$, mechanically reduces the elasticity.  Many previous studies estimate heterogeneity in $\gamma_C$ alongside a single $\alpha_C$ parameter for the population (\cite{mcrae2014infrastructure}, \cite{olmstead2009reduced}, \cite{szabo2015value}).  This paper builds on these previous studies by estimating heterogeneity in both terms to allow to for a broader range of possibilities, including that larger users may be relatively more sensitive to prices than smaller users.  To compare estimates in this paper with previous papers under non-linear pricing schedules, this approach follows the same formula for computing elasticities by (1) taking draws of $\epsilon$ and $\eta$, (2) computing predicted demand under prices that are 1\% greater than current prices, and (3) comparing the average percentage change in demand before and after the price change.  For tariff blocks equal to zero, prices are increased from 0 to 1 PhP following \cite{szabo2015value}.

@@

\begin{table}
\centering
\begin{threeparttable}
\caption{Elasticity Estimates}\label{table:elasticity}
\input{tables/elasticity_estimates} 
\begin{tablenotes}
\item {\footnotesize Elasticities are calculated from 7,151 connections making 349,265 connection-months.  Income is imputed based on demographics, locations, and their interactions.  Standard errors are bootstrapped at the connection level.}
\end{tablenotes}
\end{threeparttable}
\end{table}

The demand estimation yields an average elasticity of \input{tables/mean_elasticity}\unskip, which can be interpreted as the percentage decrease in quantity due to a one percentage increase in price.  This estimate also closely matches reduced-form elasticity estimates from price-change quasi-experiment detailed in Appendix~\ref{appendix:ratechangeelasticity}.  This estimate is in range of the 0.98 estimate from \cite{szabo2015value} as well as similar studies in the developing world that find elasticities between 0.01 and 0.81 (\cite{diakite2009proposal}, \cite{strand2005water}).

@@

Figure~\ref{figure:predictedwateruse} compares the predicted histogram of consumption with observed values.  Vertical red lines indicate the price thresholds in the tariff schedule.  By roughly matching the smoothness of the distributions especially across price thresholds, this model improves on standard models in the literature, which often predict excessive bunching at thresholds (\cite{mcrae2014infrastructure}, \cite{olmstead2009reduced}, \cite{szabo2015value}).  The model has difficultly rationalizing low levels of consumption likely because households face zero prices below this threshold.  Since preference and consumption shocks are assumed to be normally distributed, the model predicts some negative consumption values.  In the counterfactual exercises, I treat these negative realizations as zero consumption choices.  Likewise, the model also cannot account for bunching at zero (likely capturing households that go on vacation).  The model also predicts greater bunching around the upper price threshold despite little observed bunching in the data as large users tend to ignore this smaller jump in price.

@@

\begin{table}
\centering
\begin{threeparttable}
\caption{Correlation Estimates}\label{table:correstimates}
\input{tables/corr_table_group}
\begin{tablenotes}
\item { \footnotesize Errors are bootstrapped at the connection level. }
\end{tablenotes}
\end{threeparttable}
%\vspace{.3cm}
%\footnotesize{1 indexes correlations with first movers \\ while 2 indexes correlations with second movers}
\end{table}

@@

\begin{table}
\centering
\begin{threeparttable}
\caption{Hassle Cost Estimates}\label{table:step2estimates}
\input{tables/step_2_estimates}
\begin{tablenotes}
\item \footnotesize{Estimates include 548 leaking connections and 2,151 neighboring connections, making 20,299 connection-months.  \% single/home and neighbor distance are averages within wards.  Neighbor distance is in meters.   Standard errors are bootstrapped at the level of the leaking connection (and neighbors). }
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{table}
\centering
\begin{threeparttable}
\caption{Fixed Cost Estimates}\label{table:smmestimates}
\input{tables/step_3_estimates_group}
\begin{tablenotes}
\item \footnotesize{  Fixed costs are in PhP/month and prices are in PhP/m3.   \input{tables/totalhhsmm_group} households across 599 wards are included in simulated method of moment estimation using source choices.  Standard errors are bootstrapped at the connection level }
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{table}
\centering
\caption{Connection Fee Discount: Out-of-Sample Test Results}\label{table:oosestimates}
\input{tables/oos_discount_estimates_group} \\ 
\vspace{.2cm}
\footnotesize{Fixed fee and consumer surplus are in PhP/month.}
\end{table}

@@

Table~\ref{table:fixedfeediscount} summarizes the impacts of this change on source choices and consumer welfare measured in terms of the difference between current utility and utility from using the outside vendor option as in Section~\ref{section:oostest}.  While the increase in the tariff drives large users to substitute toward vendor options, the decrease in the fixed price allows some small users to join together and purchase shared connections, resulting in no change in the share of households using from a neighbor.  The third row indicates even greater substitution toward water vendors of around \input{tables/discount_vendor_change_group}percentage points in response to the higher tariff.  These effects combine to reduce average consumer surplus by \input{tables/discount_avg_surplus_change_group}PhP per month, which is over half of the initial fixed fee discount of \input{tables/c_fee_discount_group}PhP.  This decline translates into a \input{tables/discount_avg_welfare_change_group}percent loss of monthly household income.  Households in the bottom \input{tables/group_percentile_group}\unskip th percentile of water users also experience lower consumer surplus although the magnitude is much less, amounting to \input{tables/discount_low_welfare_change_group}percent of monthly household income.  Taken together, these results suggest that connection fee discount policies in this context can have unintended distributional consequences by reducing both access to piped water and welfare for the intended beneficiaries. %\footnote{Consumer surplus is measured relative to utility from using vended water.}  
% a drop of \input{tables/discount_sharing_change_group}percentage points in households using water from their neighbors The decrease in the fixed fee and increase in the tariff both act to decrease the incentives for sharing a connection and leading to no change in households using water from their neighbors. 
\begin{table}
	\centering
	\caption{Fixed Fee Discount Counterfactual}\label{table:fixedfeediscount}
	\input{tables/discount_estimates_group} \\
	\vspace{.2cm}
	\footnotesize{Low Users include the bottom 50\% of water users. \\
	Surplus and fixed fee are in PhP/month.}
\end{table}

@@

\begin{table}[!hbp]
\centering
\caption{Optimal Tariffs Without Sharing}\label{table:optimaltariffnosharing}
\resizebox{\textwidth}{!}{%
\input{tables/welfare_no_sharing_groups}
}	\\
\footnotesize{$\text{ }^{*}$ The First-Best fee does not include individual transfers imposed by the regulator. \\  Social tariffs place all welfare weight on the bottom 50\% of water users. All terms are monthly.}
\end{table}
